---
title: "Final Project"
description: |
  Using gradiant boosting to predict heart problems
author:
  - name: Colin Miner 
date: "`r Sys.Date()`"
output: distill::distill_article
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(curl)
library(caret)
library(tidyverse)
library(caret)
library(dplyr)
library(gbm)
library(psych)
```

**1. A description of your data and where you found it and what you are
predicting**

The data concerns patients who are at risk for heart disease, and
contains multiple attributes that can be risk factors or red flags for
heart disease. It was developed by combining five other publicly
available heart datasets and eliminating the duplicates. I will be using
it to predict if a patient has heart disease.

Source: <https://www.kaggle.com/fedesoriano/heart-failure-prediction>

**2. Descriptive statistics - small discussion...are the skewed or are
there outliers etc.**

The patient records are highly skewed towards men, and towards patients
with \<= 120 mg/dl fasting blood sugar. The resting BP has a high
kurtosis, partly because of one major outlier with no resting BP, and 20
records (\~2%) exceeding 180 bpm.

The average patient age is 53, which is useful for this prediction as
young people rarely have detectable heart problems on average.

The cholesterol levels vary pretty widely, and I predict will be an
important variable.

But overall the dataset is very clean.

```{r stats, warning=FALSE, message=FALSE}
heart <- read.csv("data/Heart.csv",sep=',')
psych::describe(heart)
```

**3. Describe your machine learning model (short description)**

Gradient boosting is very similar to random forests, in that it uses
many decision trees to develop a better prediction model. It differs in
how the trees are created and how the results are combined. The decision
trees are developed iteratively with each new tree helping to fix the
errors of the previous tree, to form an optimal model.

**4. Split data, run model, and find the results**

```{r gb, warning=FALSE, message=FALSE}

heart<-heart%>%
  mutate(prediction=if_else(HeartDisease==1,"YES","NO"))

set.seed(1)
#lets split the data 60/40


trainIndex <- createDataPartition(heart$HeartDisease, p = .6, list = FALSE, times = 1)

#grab the data
heartTrain <- heart[ trainIndex,]
heartTest  <- heart[-trainIndex,]

set.seed(1)

heartGBM<- train(
  form = factor(HeartDisease)~factor(Age)+factor(Sex),
  data = heartTrain,
  #here we add classProbs because we want probs
  trControl = trainControl(method = "cv", number = 10,
                           classProbs =  TRUE),
  method = "gbm",
  tuneLength = 30,
  verbose=FALSE)

tipsGBM_Pred<-predict(tipsGBM,tipsTest,type="prob")
tipsGBMtestpred<-cbind(tipsGBM_Pred,tipsTest)

tipsGBMtestpred<-tipsGBMtestpred%>%
  mutate(prediction=ifelse(Male>=.5,"Male","Female"))

confusion_obj<-confusionMatrix(factor(tipsGBMtestpred$prediction),factor(tipsGBMtestpred$sex))

ggplot(as.data.frame(confusion_obj$table))+ 
  geom_raster(aes(x=Reference, y=Prediction, fill=Freq)) + 
  geom_text(aes(x=Reference, y=Prediction, label=Freq)) +
   scale_fill_gradient2( low = "darkred", high = "pink", na.value="black", name = "Freq" )+
  scale_x_discrete(name="Actual Class") + 
  scale_y_discrete(name="Predicted Class")+
  ggtitle("Confusion is fun")+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'))


#roc
rocobj<-pROC::roc( factor(tipsGBMtestpred$sex),tipsGBMtestpred$Male)

rocobj
plot(rocobj)


#sick gains bro
pred<-ROCR::prediction(tipsGBMtestpred$Male ,factor(tipsGBMtestpred$sex))

gain <- ROCR::performance(pred, "tpr", "rpp")

plot(gain, main = "Gain Chart")


#lift
perf <- ROCR::performance(pred,"lift","rpp")

plot(perf, main="Lift curve") 

V<-caret::varImp(tipsGBM, n.trees=500)$importance%>%
  arrange(desc(Overall))

knitr::kable(V)

ggplot2::ggplot(V, aes(x=reorder(rownames(V),Overall), y=Overall)) +
geom_point( color="blue", size=4, alpha=0.6)+
geom_segment( aes(x=rownames(V), xend=rownames(V), y=0, yend=Overall), 
color='skyblue') +
xlab('Variable')+
ylab('Overall Importance')+
theme_light() +
coord_flip() 

```

**5. Describe your results. How can this be useful**
